---
title: "NHL Goalie Evaluation"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    css: style.css
    includes:
      before_body: header.html
author: Ange Michaella Niyonkuru (8962161), Erhardt Jansen van Rensburg (8150118), Maia Pelletier (8245669)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, fig.align = 'center', fig.height = 8)

## Load libraries
library(tidyverse)
library(teamcolors)
library(ggsci)
library(rnaturalearth)
library(sf)
library(sp)
library(fishualize)
library(plotly)
library(ggrepel)
library(patchwork)
library(janitor)
library(readxl)
library(lubridate)
library(ggthemes)
library(stringi)
library(FSelector)
library(kableExtra)
library(broom)

theme_set(theme_minimal())

path <- 'data/'

# Combine MetaHockey ------------------------------------------------------

mh_files <- paste0(path, list.files(path, pattern = 'MH_'))

## General cleaning function
clean_mh_data <- function(mh_data) {
  
  tidy_mh_data <- 
    mh_data %>%
    clean_names() %>%
    mutate(cap_hit = as.numeric(cap_hit)) %>%
    mutate(dob = date(parse_date_time(dob, orders = c('mdy', 'ymd'))))
  
  if (any(str_detect(names(tidy_mh_data), 'chip'))) {
    tidy_mh_data <- 
      tidy_mh_data %>%
      mutate(chip = as.numeric(chip))
  }
  
  if(any(str_detect(names(tidy_mh_data), 'dist'))) {
    tidy_mh_data <- 
      tidy_mh_data %>%
      mutate(dist = as.character(dist))
  }
  
  return(tidy_mh_data)
}


mh_szns <- paste0(12:17, '-', 13:18)

## Read/clean data and append to list
mh_data_list <- list()
for (i in 1:length(mh_files)) {
  mh_data_list[[i]] <- 
    read_excel(mh_files[i]) %>%
    clean_mh_data() %>%
    mutate(szn = mh_szns[i])
}

## Dataset specific cleaning
mh_data_list[[2]] <- 
  mh_data_list[[2]] %>% 
  rename(`gp` = gp_16, gs = gs_17, sa = sa_21, sv_percent = sv_percent_24)

mh_data_list[[3]] <- 
  mh_data_list[[3]] %>%
  rename(gp = gp_17)

## Combine data into one dataset
new_mh_data <- tibble()
for (i in 1:length(mh_data_list)) {
  new_mh_data <- bind_rows(new_mh_data, mh_data_list[[i]])
}

## Clean dataset
getblanks <- function(blanks) {
  
  blank <- sum(is.na(new_mh_data[blanks]))
  column_name <- colnames(new_mh_data[blanks])
  ans <- tibble(column_id = blanks, column_name = column_name,  NA_amount = blank)
  
  return(ans)
}

## Find number of blanks
response <- tibble()
for (i in 1:ncol(new_mh_data)) {
  response  <- bind_rows(response, getblanks(i))
}

## Remove column if filled with blanks (no match in other datasets)
for (y in 1:nrow(response)) {
  
  if (response[y, "NA_amount"] > 81) {
    
    name_column <- paste((response[y,"column_name"]))
    
    new_mh_data <- 
      new_mh_data %>% 
      select(-name_column)
  }
}

mh_data <- 
  new_mh_data %>%
    unite('player', c('first_name', 'last_name'), sep = ' ')

# Combine HockeyReference -------------------------------------------------

hr_files <- paste0(path, list.files(path, pattern = 'HR_'))

num_of_sheets <- length(excel_sheets(hr_files))

hr_szns <- paste0(12:18, '-', 13:19)

# General cleaning function
clean_hr_data <- function(hr_data) {
  
  hr_data <- 
    hr_data %>%
    clean_names() %>%
    mutate(szn = hr_szns[i]) %>%
    mutate(player = str_remove(player, '\\\\[\\w]+')) %>%
    mutate(player = str_remove(player, '\\*')) %>%
    add_count(player, name = 'num_of_teams') %>%
    mutate(
      num_of_teams = ifelse(num_of_teams > 1, num_of_teams - 1, num_of_teams)
    )
  
  clean_data <-
    bind_rows(
      ## Traded players
      hr_data %>%
        filter(num_of_teams > 1) %>%
        filter(tm == 'TOT'),
      ## Players that played on the same team all szn
      hr_data %>%
        filter(num_of_teams == 1)
    ) %>%
    rename(team = tm) %>%
    arrange(rk)
  
  
  return(clean_data)
}

## Append data from each sheet in excel file to empty list and clean
hr_data_list <- list()
for (i in 1:num_of_sheets) {
  hr_data_list[[i]] <- 
    read_excel(hr_files, sheet = i, skip = 1) %>% 
    clean_hr_data()
}

## Combine data into on dataset
hr_data <- tibble()
for (i in 1:length(hr_data_list)) {
  hr_data <- 
    bind_rows(hr_data, hr_data_list[[i]])
}


# Combine MoneyPuck -------------------------------------------------------

mp_files <- paste0(path, list.files(path, pattern = 'MP_'))

hr_szns <- paste0(12:18, '-', 13:19)

mp_data_list <- list()
for(i in 1:length(mp_files)) {
  mp_data_list[[i]] <- 
    read_csv(mp_files[i]) %>%
      clean_names() %>%
      mutate(szn = hr_szns[i])
}

mp_data <- tibble()
for (i in 1:length(mp_data_list)) {
  mp_data <- bind_rows(mp_data, mp_data_list[[i]])
}

# Combine CapFriendly -------------------------------------------------

cf_files <- paste0(path, list.files(path, pattern = 'CF-'))

num_of_sheets <- length(excel_sheets(cf_files))

cf_szns <- excel_sheets(cf_files)

## Append data from each sheet in excel file to empty list
cf_data_list <- list()
for (i in 1:num_of_sheets) {
  cf_data_list[[i]] <- 
    read_excel(cf_files, sheet = i) %>% 
    clean_names() %>%
    mutate(szn = cf_szns[i])
}

## Small tidying
for (i in 1:length(cf_data_list)) {
  cf_data_list[[i]] <-
    cf_data_list[[i]] %>%
    # When a player dies(during the szn?), it denotes "Died" in Age column
    mutate(age = str_extract(age, '\\d+'),
           age = as.numeric(age)) %>%
    # The players column contains a numbered index (i.e. 1. Carey Price)
    mutate(player = str_remove(player, '[0-9]*\\.\\s*'))
}
  
## Combine data into on dataset
cf_data <- tibble()
for (i in 1:length(cf_data_list)) {
  cf_data <- 
    bind_rows(cf_data, cf_data_list[[i]]) 
}

cf_data <-
  cf_data %>%
    separate(drafted, into = c('overall', 'draft_round', 'draft_year'), sep = ' - ') %>%
    separate(draft_year, into = c('draft_year', 'draft_team'), sep = ' ') %>%
    mutate(overall = as.numeric(overall),
           draft_year = as.numeric(draft_year),
           draft_team = str_remove_all(draft_team, '[\\( \\)]'))

# Combine NHL Standings ---------------------------------------------------

standings_files <- paste0(path, list.files(path, pattern = 'nhl_standings'))

num_of_sheets <- length(excel_sheets(standings_files))

standings_szns <- paste0(12:18, '-', 13:19)

## Combine data into one dataset
standings_data <- tibble()
for (i in 1:num_of_sheets) {
  add <- 
    read_excel(standings_files, sheet = i, skip = 1) %>% 
    clean_names() %>%
    mutate(szn = standings_szns[i]) %>%
    select(rk, team = x2, w, l, ol, pts, szn) %>%
    mutate(team = str_remove(team, '\\*'))
  
  standings_data <- bind_rows(standings_data, add)
}

```

## Introduction
***

NHL goalies are kind of an enigma. They're notoriously hard to evaluate and predict in a sport that is already so chaotic. They are also incredibly important to their teams and often you will hear hockey analysts citing teams' goalies as the reason that the team is preforming well or poorly.  

We are interested specifically in evaluating NHL goalies for the purpose of understanding their contracts. How do teams decide how much to pay their goalies? We want to inspect goalies in the post-lockout seasons and take a look at the best, the worst and how much they get paid.  

## Objective
***
The objective of this report is to evaluate NHL goalies and their contracts in the post-lockout era and then build a model to try and predict how much goalies should be getting paid.  

## NHL Goalies 
***

We wanted to specifically preform this analysis during the post-lockout era (2013-present). Hockey has changed so much in these seasons, comparing goalies from post-lockout era hockey to pre-lockout doesn't quite make sense.  

### Overview of Goalies  

How many unique goalies played per season?

```{r unique gp per szn}
hr_data %>%
  distinct(player, szn) %>%
  count(szn, name = 'Number of goalies') %>%
  rename(Season = szn) %>%
  kable() %>%
  kable_styling(bootstrap_options = c('hover'), full_width = F) %>%
  row_spec(0, background = "#e6f0ff")
```

We see that ~95 goalies played each season. This may seem like a lot but this actually makes sense. From 2013-2017 there were 30 NHL teams in the league and from 2017-2019 there were 31. Each NHL team usually has a starting goalie, a back-up goalie, and a third-string "emergency" goalie. Teams calling-up goalies from their farm teams in emergency situations is also common. So you would assume that **number of teams in league * 3** is the approximate number of goalies who play each season.  

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

<b>Fact:</b> In 2018-19, between goalies getting constantly blown out and injured, the Philadelphia Flyers were infamous for their goaltending trouble. They iced <i>8 goalies</i> during the regular season, the highest of any team post-lockout.

</div>

```{r philly goalies}
read_csv('data/flyers_goaliedata_19.csv') %>%
  clean_names() %>%
  arrange(desc(gp)) %>%
  select(Player = player, Team = tm, `Games played` = gp, `Save %` = sv_percent) %>%
  kable() %>%
  kable_styling(bootstrap_options = c('condensed', 'hover'), full_width = F)
```


In total there were `r hr_data %>% distinct(player) %>% nrow()` goalies that played at least 1 game in the NHL from 2013-2019. However, it makes no sense to evaluate a goalie based on their performance in just a few games because of the variation between events occuring in each hockey game. One of the many mysteries surrounding goalies is after how many games can we draw definitive conclusions about their quality of play. Let's look at how many games goalies were playing per season and over their whole careers.  

```{r histograms, fig.height=5}
gp_szn <-
  ggplot(hr_data, aes(gp)) +
  geom_histogram(binwidth = 5, fill = '#a7d2cb', alpha = 0.8) +
  labs(x = 'Games played per season by goalies', y = 'Frequency') +
  theme(plot.caption = element_text(color = 'grey50'))

gp_overall <-
  hr_data %>%
  group_by(player) %>%
  summarise(gp = sum(gp)) %>%
  ggplot(aes(gp)) +
  geom_histogram(binwidth = 25, fill = 'thistle', alpha = 0.8) +
  labs(x = 'Games played overall by goalies', y = 'Frequency',
       caption = '1 Game played = 1 or more minute played on ice') +
  theme(plot.caption = element_text(color = 'grey50'))

gp_szn + gp_overall
```

In the case of this analysis, we were interested specifically in goalies who had played 40 or more games in the NHL from 2013-2019. Our reasoning behind this is that a full NHL season is 82 games so 40 games is about half a "season" played and seems like a reasonably large enough sample to evaluate a goalie's play. The one downside to this evaluation is that it will devalue the contributions of back-up goalies, but back-up goalies usually don't last as long as starters anyway and thus, their contracts tend to stay around league minimum.  

### Goalie demographics  

It is well-known that most hockey players origniate from North America, Russia, or Scandinavia. Is this true for goalies specifically?  

```{r map, fig.height=6, fig.width=9}
goalie_countries <- 
  cf_data %>%
    count(country, name = "num_of_goalies") %>%
    na.omit()

world <- 
  ne_countries(
    scale = "medium", 
    returnclass = "sf"
  )

world %>%
  left_join(goalie_countries, by = c('name' = 'country')) %>%
  ggplot() +
  geom_sf(aes(fill = num_of_goalies)) +
  labs(fill = "Number of goalies") + 
  scale_fill_fish(option = 'Scarus_globiceps') +
  theme_void() +
  theme(legend.position = "bottom",
        legend.key.height = unit(2, 'mm'),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8))
```

We can see that this holds true for goalies for the most part, but it's interesting to note that there are goalies who originate from Denmark and even the UK.  

The old hockey idea is that goalies are supposed to be **big**. The taller and wider they are, the more space they take up in the net. However, with the evolution of hockey skill came the need for more athletic goalies with lightning reflexes. The butterfly style, now commonplace in the NHL, requires goalies to be flexible. So what do goalies look like in the NHL now?  

```{r, warning=F, message=F, fig.height=7}
cf_data_player_height_weight <- 
  cf_data %>%
  select(player, weight, height, age, country) %>%
  mutate(
    weight = parse_number(sub(".*-","", weight)), 
    height = parse_number(sub(".*-","",height))
    ) %>%
  mutate(age = case_when(
    age < 20 ~ 'Under 20',
    age >= 20 & age < 25 ~ '20-24',
    age >= 25 & age < 30 ~ '25-29',
    age >= 30 & age < 35 ~ '30-34',
    age >= 35 & age < 40 ~ '35-39',
    age > 40 ~ 'Over 40',
  ))

p <-
  cf_data_player_height_weight %>%
  ggplot(aes(height, weight)) +
  geom_jitter(aes(color = age, label = player, label2 = country), 
              alpha = 0.65, size = 3) +
  labs(x = 'Goalies\'s Height (cm)', 
       y = 'Goalies\'s Weight (kg)', 
       color = 'Age') +
  scale_color_fish_d(option = 'Callanthias_australis')

ggplotly(p)
```

So we can see that the largest chunck of goalies seems to fall between 183 - 193cm (6'0 - 6'3 ft) in height and 85 - 95kg (187 - 210lbs) in weight. So only slightly taller and heavier than the average adult male, which is what you would expect from a professional athelete but is not quite what you would expect if you tend to think of goalies as big players.  

### The Best (and Worst) of the NHL
We wanted to see some how much the best and the worst goalies are getting paid.  This is to see if they are getting over paid becuse they are not preforming well, or to see if a team is getting a good deal by not paying too much for a really goog goalie.   

We used four different metrics to determine how good a goalie is. The first is **save percentage** which is the percent of shots the goalie allows in compared to all shots faced. The second is **goals against average**, which is the average of goals that the goalie let in during a game. The third is *percentage of how many games the goalie won*. The final metric is **goals saved above average** which shows the average goals saved compared to the league average.

#### The top 10 averages by goalies over the past 6 seasons  

Let's find out who the top 10 goalies were over the past 6 seasons. 


```{r top 10 avgs}

hr_data <- hr_data %>% mutate(w_percent = w/gp)

Salaries <- cf_data %>% 
  group_by(player) %>% 
  summarise(mean_aav = mean(aav)) %>% 
  mutate(player = replace(player,player == "Marc-André Fleury","Marc-Andre Fleury")) %>%
  mutate(player = replace(player,player == "Jaroslav Halák","Jaroslav Halak")) %>% 
  mutate(player = replace(player, player == "Eddie Läck","Eddie Lack")) %>% 
  mutate(player = replace(player, player == "Jacob Markström","Jacob Markstrom")) %>%
  mutate(player = replace(player, player == "Petr Mrázek","Petr Mrazek"))

Best_Goalies <- 
  hr_data %>% 
  filter(!(player == "Martin Jones" & szn == "14-15")) %>%
  group_by(player) %>% 
  summarise(gp = sum(gp),
            w = sum(w),
            ga = sum(ga),
            sa = sum(sa),
            sv = sum(sv),
            ) %>%
  filter(gp > 120) %>% 
  mutate(mean_w_percent = w/gp, 
         mean_sv_percent = sv/sa,
         mean_gaa = ga/gp,
         avg_sv = mean(mean_sv_percent),
         mean_gsaa = (sa * (1-avg_sv))- ga,
         )

Best_Goalies <- Best_Goalies %>% 
  left_join(Salaries, by = c("player" = "player"))

Mean_sv_plot <- 
  Best_Goalies %>% 
  top_n(10, mean_sv_percent) %>%
  ggplot(aes(x = mean_sv_percent, y = mean_aav)) + 
  geom_point(size = 2, color = '#a7d2cb') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Save Percentage", y = "Average Annual Salary")+
  scale_y_continuous(labels = scales::dollar)

Mean_gaa_plot <- 
  Best_Goalies %>% 
  top_n(-10, mean_gaa) %>%
  ggplot(aes(x = mean_gaa, y = mean_aav)) + 
  geom_point(size = 2, color = '#f2d388') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Against Average", y = "Average Annual Salary")+
  scale_y_continuous(labels = scales::dollar)

mean_w_percent_plot <- 
  Best_Goalies %>% 
  top_n(10, mean_w_percent) %>% 
  ggplot(aes(x = mean_w_percent, y = mean_aav)) + 
  geom_point(size = 2, color = '#c98474') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Percentage of Games Won", y = "Average Annual Salary")+
  scale_y_continuous(labels = scales::dollar)

mean_gsaa_plot <- 
  Best_Goalies %>% 
  top_n(10, mean_gsaa) %>% 
  ggplot(aes(x = mean_gsaa,y = mean_aav))+ 
  geom_point(size = 2, color = '#874c62') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Saved Above Average", y = "Average Annual Salary")+
  scale_y_continuous(labels = scales::dollar)

(Mean_sv_plot | Mean_gaa_plot) / (mean_w_percent_plot | mean_gsaa_plot)
```

From the graphs we can see some common names pop up between the four metrics, for example Carey Price, Sergei Bobrovsky are all names really good goalies and that is why they get paid around 7 Million per year on average. 

Some other good goalies that are making around 3 Million per year are Matt Murray, Andrei Vasilevskiy. Those teams are getting a good deal for top quality goaltending. 


#### Top goalies by season 

We then just wanted to see what the top goalie was in each metric for any particular season.  

```{r top G by szn}
# Finding the best goalies by Season
# creates tables for each stat we are looking at

top_goalie_sv <- function(season){
  top_goalie_stats <- hr_data %>% filter(szn == season, gp > 40) 
  top_goalie_sv_percent <- top_goalie_stats %>% top_n(1,sv_percent) %>%
    select(player, sv_percent, szn)
  return(top_goalie_sv_percent)
}

top_goalie_gaa <- function(season){
  top_goalie_stats <- hr_data %>% 
    filter(szn == season, gp > 40)
  goalie_gaa <- top_goalie_stats %>% 
    top_n(-1,gaa) %>% 
    select(player, gaa, szn)
  return(goalie_gaa)
}


top_goalie_w_percent <- function(season){
  top_goalie_stats <- hr_data %>% filter(szn == season, gp > 40)
  goalie_w_percent <- top_goalie_stats %>% top_n(1,w_percent) %>% 
    select(player, w_percent, szn)
  return(goalie_w_percent)
}


top_goalie_gsaa <- function(season){
  top_goalie_stats <- hr_data %>% filter(szn == season, gp > 40)
  goalie_gsaa <- top_goalie_stats %>% top_n(1,gsaa) %>% 
    select(player, gsaa, szn)
  return(goalie_gsaa)
}

#Loop to fill in the tables

best_sv_year <- tibble()
best_gaa_year <- tibble()
best_win_percent_year <- tibble()
best_gsaa_year <- tibble()
for (i in 1:7){
  best_sv_year <- bind_rows(best_sv_year,top_goalie_sv(hr_szns[i]))
  best_gaa_year <- bind_rows(best_gaa_year,top_goalie_gaa(hr_szns[i]))
  best_win_percent_year <- bind_rows(best_win_percent_year,top_goalie_w_percent(hr_szns[i]))
  best_gsaa_year <- bind_rows(best_gsaa_year,top_goalie_gsaa(hr_szns[i]))
}



#Plots to see best goalies by season 


sv_plot <- 
  best_sv_year %>% 
  ggplot(aes(x = szn, y = sv_percent)) + 
  geom_point(size = 2, color = '#a7d2cb') +
  geom_text_repel(aes(label = player), size = 3) +
  coord_flip() + 
  labs(x = "Season", y = "Save Percentage")

gaa_plot <- 
  best_gaa_year %>% 
  ggplot(aes(x =szn, y = gaa)) + 
  geom_point(size = 2, color = '#f2d388') +
  geom_text_repel(aes(label = player), size = 3) +
  coord_flip() + 
  labs(x = "Season", y = "Goals Against Average")

w_percent_plot <- 
  best_win_percent_year %>% 
  ggplot(aes(x = szn, y = w_percent)) + 
  geom_point(size = 2, color = '#c98474') + 
  geom_text_repel(aes(label = player), size = 3) +
  coord_flip() + 
  labs(x = "Season", y = "Win Percentage")

gsaa_plot <- 
  best_gsaa_year %>% 
  ggplot(aes(x = szn, y = gsaa)) + 
  geom_point(size = 2, color = '#874c62') + 
  geom_text_repel(aes(label = player), size = 3) +
  coord_flip() + 
  labs(x = "Season", y = "Goals Saved Above Average")

year_plots <- (sv_plot | gaa_plot) / (w_percent_plot | gsaa_plot)

year_plots 
```
From here we see some of the names again that appeared on the top averages graph. Notibly Carey Price and Sergei Bobrovsky.  We also want to point out that many of the top goalies per season were with the team that either won the Stanely Cup or are with a team that was in the top of the standings. This will be shown later when we see who the goalies were for the top preforming teams. 

#### Worst Goalies over past seasons
Now that we took a look at the best performing goalies, it is only fair that we analyze the 10 worst performing goalies over the past 6 seasons.  

```{r worst G over past szns}

#Worst Goalies by averages overall

worst_mean_sv_plot <- Best_Goalies %>% 
  top_n(-10,mean_sv_percent) %>%
  ggplot(aes(x = mean_sv_percent, y = mean_aav)) + 
  geom_point(size = 2, color = '#a7d2cb') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Save Percentage", y = "Average Annual Salary") +
  scale_y_continuous(labels = scales::dollar)

worst_mean_gaa_plot <- Best_Goalies %>% 
  top_n(10,mean_gaa) %>%
  ggplot(aes(x = mean_gaa, y = mean_aav)) + 
  geom_point(size = 2, color = '#f2d388') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Against Average", y = "Average Annual Salary")+
  scale_y_continuous(labels = scales::dollar)

worst_mean_w_percent_plot <- Best_Goalies %>% 
  top_n(-10, mean_w_percent) %>% 
  ggplot(aes(x = mean_w_percent, y = mean_aav)) + 
  geom_point(size = 2, color = '#c98474') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Percentage of Games Won", y = "Average Annual Salary")+
  scale_y_continuous(labels = scales::dollar)


worst_mean_gsaa_plot <- Best_Goalies %>% 
  top_n(-10, mean_gsaa) %>% 
  ggplot(aes(x = mean_gsaa,y = mean_aav))+ 
  geom_point(size = 2, color = '#874c62') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Saved Above Average", y = "Average Annual Salary")+
  scale_y_continuous(labels = scales::dollar)


Worst_average_plots <- (worst_mean_sv_plot | worst_mean_gaa_plot) / (worst_mean_w_percent_plot | worst_mean_gsaa_plot)

Worst_average_plots 
```
There is one goalie to pay attention too on this chart, and that is Cam Ward.  He is one of the worst performing goalies on average over the past 6 seasons with one the highest salaries that is around $5 Million a year. He has the lowest Save percentage and the worst goals saved above average with a -70 goals saved. Interesting to note that that Cam Ward carried his team to a Stanley Cup however in 2006 while posting a 0.882 save percentage.  

Mike Smith is another high paid goalie that is constantly bad with his goals against average and games won. The other goalies are all pad a lower amount that is around $2M per year, which means that their teams luckily are not spending too much money on a bad goalie. 


#### Worst goalies over last 6 seasons

We wanted to see who were some of the truly worst goalies in the past 6 seasons.  A truly bad goalie has a save percentage of less than 0.9 in the season and has played more than 40 games. 

```{r, warning = FALSE, message = FALSE, fig.height=9, fig.width=9}

# Some really bad goalies by year 

missing_salary <- 
  read_xls("data/MH_nhl_goalies_2017-2018.xls") %>% 
  clean_names() %>% 
  unite('player', c('first_name', 'last_name'), sep = ' ') %>% 
  filter(player == "Scott Darling") %>% 
  select(player, salary)

worst_goalies <- hr_data %>%
  filter(gp > 40 & sv_percent < 0.9)

worst_salaries <- cf_data %>% select(player, aav, szn)

worst_goalies <- worst_goalies %>% 
  left_join(worst_salaries,by = c("player" = "player", "szn" = "szn")) %>%
  left_join(missing_salary, by = c("player" = "player")) %>% 
  mutate(aav = replace_na(aav,0)) %>% 
  mutate(salary = replace_na(salary,0)) %>% 
  mutate(aav = aav+ salary) %>% 
  unite('player_szn', c('player','szn'), sep = ' | Season:', remove = FALSE) %>%
  select(-salary) %>%
  mutate(w_percent = w/gp)

wg_sv_percent <- worst_goalies %>% 
  ggplot(aes(x = sv_percent, y = aav, color = szn)) + 
  geom_point(size = 2) +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Save Percentage", y = "Annual Average Salary", color = "Season") +
  scale_color_fish_d(option = 'Scarus_globiceps') +
  scale_y_continuous(labels = scales::dollar)

wg_gsaa <- worst_goalies %>% 
  ggplot(aes(x = gsaa, y = aav, color = szn)) + 
  geom_point(size = 2) +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Saved Above Average", y = "Annual Average Salary", color = "Season")+
  scale_color_fish_d(option = 'Scarus_globiceps')+
  scale_y_continuous(labels = scales::dollar)

wg_gaa <- worst_goalies %>% 
  ggplot(aes(x = gaa, y = aav, color = szn)) + 
  geom_point(size = 2) + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Against Average", y = "Annual Average Salary", color = "Season") +
  scale_color_fish_d(option = 'Scarus_globiceps')+
  scale_y_continuous(labels = scales::dollar)

wg_w_percent <- worst_goalies %>% 
  ggplot(aes(x = w_percent, y = aav, color = szn)) + 
  geom_point(size = 2) + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Win Percentage", y = "Annual Average Salary", color = "Season") +
  scale_color_fish_d(option = 'Scarus_globiceps')+
  scale_y_continuous(labels = scales::dollar)

Worst_goalies_plot <- (wg_sv_percent | wg_gaa) / (wg_w_percent | wg_gsaa)
Worst_goalies_plot + plot_annotation(
   caption = 'A bad season has a save percentage of less than 0.90'
 )

```

From the charts we see that there are only 8 goalies who had a really bad season since the lockout.  Most of them got paid over 4M for that year.  Some notable mentions include Johnathan Quick who had his worst season in 2018-2019.  He is a declining performing goalie who once led his team to two Stanley Cups in his prime. Martin Jones had a bad season in 2018-2019 as well and his team lost in the conference finals in the Stanley Cup.  Roberto Luongo was another great goalie back in 2010 who led team Canada to a Gold medal in the Vancouver Olympics and led the Vancouver Canucks to game 7 of the Stanely Cup final. 

Mike Smith is getting paid way to much since he has no track record like some of the other goalies in this category. He really should not be getting paid what he is since the the other goalies mentioned above were on long term contracts and are nearing the end of their carreer. 

#### First place team goalie performance

A phenonmeon surrounding goalies is that a goalie preforming well generally correlates with their team preforming well and vice versa. One has to wonder if the goalie has this effect on their team (i.e. implying that a good goalie can make a team be good) or if a good team inflates a goalies perceived quality (i.e. implying that a good team makes a good goalie). Obviously there are other factors that could be considered in team and goalie quality but to isolate this question a bit, let's investigate the performance of the goalies belonging to the first place team in the league (most points at the end of the regular season) over the last couple seasons.  

###### The triangle point is league average.  

```{r first place team goalies, message=FALSE, warning=FALSE, fig.height=9}
league_avgs <-
  hr_data %>%
  filter(gp > 20) %>%
  group_by(szn) %>%
  summarise(
    avg_sv_percent = mean(sv_percent),
    avg_gaa = mean(gaa)
  ) %>%
  ungroup()

starting_goalies_best <- 
  standings_data %>%
  filter(rk == 1) %>%
  select(-w, -l, -ol, -pts) %>%
  left_join(read_csv('data/NHLTeams.csv')) %>%
  left_join(hr_data, by = c('abbrev' = 'team', 'szn' = 'szn')) %>%
  group_by(team, szn) %>%
  top_n(1, gp) %>%
  select(szn, player, sv_percent, gaa) %>%
  ungroup()

first_place_goalies <- 
  starting_goalies_best %>%
  left_join(league_avgs, by = 'szn') %>%
  mutate(sv_percent_diff = sv_percent - avg_sv_percent,
         gaa_diff = gaa - avg_gaa)

starting_goalies_sv_percent <- 
  ggplot(first_place_goalies) +
    geom_point(aes(x = szn, y = sv_percent, color = sv_percent_diff), 
               size = 4) +
    geom_point(aes(x = szn, y = avg_sv_percent),
               show.legend = F, size = 3, color = 'grey30', shape = 17) +
    geom_text_repel(aes(x = szn, y = sv_percent, label = player), 
                    size = 3) +
    labs(y = 'Save percent', x = NULL, color = 'Differential') +
    ggtitle('League Average vs First Place save percent') +
    scale_color_fish(option = 'Scarus_globiceps') +
    coord_flip() +
    theme(legend.position = 'top',
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 8),
          legend.key.height = unit(2, 'mm'),
          plot.title = element_text(hjust = 0.5))

starting_goalies_gaa <- 
  ggplot(first_place_goalies) +
  geom_point(aes(x = szn, y = gaa, color = gaa_diff), 
             size = 4) +
  geom_point(aes(x = szn, y = avg_gaa),
             show.legend = F, size = 3, color = 'grey30', shape = 17) +
  geom_text_repel(aes(x = szn, y = gaa, label = player), 
                  size = 3) +
  labs(y = 'Goals against avg', x = NULL, color = 'Differential') +
  ggtitle('League Average vs First Place GAA') +
  scale_color_fish(option = 'Scarus_globiceps', direction = -1) +
  coord_flip() +
  theme(legend.position = 'top',
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.key.height = unit(2, 'mm'),
        plot.title = element_text(hjust = 0.5))

starting_goalies_comb <- starting_goalies_sv_percent / starting_goalies_gaa
starting_goalies_comb
```
A lot of these goalies were goalies that showed up in the section about best goalies in the NHL, and we can see that first place teams got some truly great goaltending, outpreforming league average by quite a bit each season. We could draw the conclusion that a good goalie tends to have a correlation with team success.  

## Goalie Contracts
***

Now, we will dive into goalie contracts. How much have goalies been getting paid in the last 6 seasons and how has that been changing? Have contract values been going up or down? Who gets paid the most and do they work hard for the money?  

When analyzing goalie contracts, most emphasis was put on analyzing the *average annual value* (AAV) of the contract (i.e. how much they make per year). The way contracts work in the NHL (and most professional sports with a salary cap) can be confusing. Some contracts are front-heavy (get paid the majority of the contract in the first few years), some are padded with bonuses from the signing teams, some have salary retention from previous teams. All of this is done to try and circumvent the salary cap, on a team management level. Thus, we feel it's safe to analyze goalies using their AAV since using actual salary would be hard to understand, interpret, and every salary varies from goalie to goalie and team to team. At the end of the day, every goalie gets the full-value of their contract and its easy to understand and interepret as if they got paid an equal amount of their contract every season.  

### How much goalies get paid

So on average how much has an NHL goaltender, who has played more than 40 NHL games, gotten paid?

```{r aav distribution by szn, fig.height=5}
aav_gp_data <-
  cf_data %>%
  select(player, aav, szn) %>%
  left_join(
    cf_data %>%
      group_by(player) %>%
      summarise(gp = sum(gp)), 
    by = 'player'
  ) %>%
  filter(gp > 40)

aav_gp_data %>%
  mutate(szn = paste0(20, szn)) %>%
  ggplot(aes(szn, aav, color = szn, fill = szn)) +
  geom_violin(alpha = 0.8,
              show.legend = F, trim = T, width = 0.9) +
  geom_boxplot(width = 0.1, color = 'grey90', fill = 'grey50', alpha = 0.1) +
  labs(x = NULL, y = 'Average annual value ($)') +
  #ggtitle('Distribution of goalie average annual value') +
  scale_y_continuous(labels = scales::dollar) +
  fishualize::scale_color_fish_d(option = 'Scarus_globiceps') +
  fishualize::scale_fill_fish_d(option = 'Scarus_globiceps') +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text.x = element_text(size = 12, color = 'grey25'),
        axis.text.y = element_text(size = 7, color = 'grey25'),
        axis.title.y = element_text(size = 9, color = 'grey25', face = 'italic'),
        plot.title = element_text(hjust = 0.5, size = 16, color = 'grey15', face= 'italic'))
```

Another way goalies are weird is that goalies in the past have not usually made as much as their skater teammates. This is starting to change however. For example, Carey Price signed a contract extension that took effect in the 18-19 season for \$10.5M per year over 8 years, \$84M in total by the time the contract is up (see above plot). This was an unprecedented signing at the time, but as teams start to value good, consistent goalies more and more, we are likely to see more high-valued goalie contracts. We can see from the plot that the median AAV has shifted in the past two seasons from \$2M to closer to \$3M. Also the AAV distribution is more spread out in the last couple seasons, instead of being clumped all near the lower end.  

### Contract inflation

We saw from the previous section that the median AAV has risen about a \$1M in the last two seasons. In general, players are signing much bigger contracts in these recent years than before. Contracts are inflating such that where a \$6M contract for a goalie used to be *huge*, it might become standard for starting goaltenders in the next couple years to be making around $6M AAV.  

To investigate contract inflation, we can look at the median AAV in each expiry year of current signed contracts. 

```{r Contract inflation, warning=FALSE, message=FALSE, fig.height=5}
cf_data %>%
  select(exp_year, aav) %>% 
  group_by(exp_year) %>%
  mutate(med_aav = median(aav)) %>%
  ggplot(aes(exp_year, med_aav)) +
  geom_step(color = 'turquoise3', size = 1) +
  geom_rug(position = 'jitter', color = 'grey30', alpha = 0.8, sides = 'right') +
  labs(x = 'Contract expiry year', y = 'Median average annual value of contract') +
  scale_y_continuous(labels = scales::dollar) +
  theme_classic() +
  theme(axis.line = element_blank())
```

The plot proves our point about contract inflation since the median AAV upon contract expiry is rising with each year. It's important to note the data density at each year (represented by the clustered dashes on the opposite axes). There are much less contracts signed past 2023 right now and therefore the median is only being calculated between 1-2 star player contracts (Carey Price at $10.5M). The dramatic increase past 2024 will be dampered a bit by goalies signing smaller deals that expire in past 2024.  

### Top paid players post-lockout

So do goalies "deserve" the money they get? As we pointed out before, it's hard to know whether a good goalie makes a good team or vice versa. But if we put stock into the idea that the goalie might be the most important player on a team, then the ones who are getting paid well, should be *really* good.  

```{r top 10 paid, warning = FALSE, fig.height=6}

## Get top 10 paid players in league over last 7 years
top_10_paid <- 
  cf_data %>%
    group_by(player) %>%
    top_n(1, aav) %>%
    ungroup() %>%
    distinct(player, aav, team) %>%
    top_n(10, aav)

## Plot top 10 paid players
top_10_paid %>%
  mutate(team = substr(team, start = str_length(team) - 2, stop = str_length(team))) %>%
  mutate(player = paste0(player, '\n(', team, ')     ')) %>%
  ggplot() +
  geom_segment(aes(x = reorder(player, aav), xend = player, y = 0, yend = aav), color = 'grey30') +
  geom_point(aes(x = player, y = aav, color = aav), size = 5, show.legend = F) +
  labs(x = NULL, y = 'Average annual value') +
  coord_flip() +
  scale_color_fish(option = 'Scarus_globiceps') +
  scale_y_continuous(labels = scales::dollar) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text.y = element_text(size = 10, color = 'grey25'),
        plot.title = element_text(hjust = 0.5, size = 16, color = 'grey25', face= 'italic'))

```

From the section on the Best and Worst goalies of the NHL, this top 10 most paid is actually very interesting. For example, we saw goalies *Carey Price*, *Henrik Lundqvist*, *Sergei Bobrovsky*, *Tukka Rask*, and *Corey Crawford* all show up in our look at the best goalies in the NHL. While *Braden Holtby* has never really posted elite numbers, he has been the goalie for the first place team (Washington Capitals) twice in the last 6 seasons and won the Stanley cup with them in 17-18. We would say that these goalies deserve to be in the top 10 most paid and seem to actually work hard for the money.  

However, we saw Cam Ward and Corey Schneider show up in our worst goalies in the NHL sections. They have not posted even mediocre numbers in the last 6 seasons and yet continue to be very well paid. We would consider these bad-value contracts for these teams.    


## How Much Should Goalies Be Making  
***

We are also interested to see if we can predict (or correct) goalie contracts. We collected goalie statistics and data from various sources including advanced goalie statistics to create a multivariate linear regression model. Thus, given the input data, we will have an idea of how much a goalie should be making.  

The first step was to combine all our data sources and take a look at our possible predictors. While combining and prepping the data, we also remove the data from goalies who we want to run through the model once it's built.  

```{r Regression prep}
## Prep and combine data

cf_reg_data <- 
  cf_data %>%
  mutate(
    ## Removes accents off names
    player = stri_trans_general(player, 'latin-ascii'),
    ## Transform all names to lowercase
    player = str_to_lower(player)) %>%
    ## Remove redundant cols
    select(-weight, -height, -pos, -team, -age, -gp, -w, -l, -so, -gaa, -sv_percent)

hr_reg_data <-
  hr_data %>%
  ## Transform all names to lowercase
  mutate(player = str_to_lower(player)) %>%
  # Filter for post lockout szns
  filter(szn != '12-13') %>%
  # Fix nicknames
  mutate(player = case_when(
    str_detect(player, 'cal heeter') ~ str_replace(player, 'cal heeter', 'calvin heeter'),
    str_detect(player, 'matt o\'connor') ~ str_replace(player, 'matt o\'connor', 'matthew o\'connor'),
    str_detect(player, 'eddie pasquale') ~ str_replace(player, 'eddie pasquale', 'edward pasquale'),
    TRUE ~ as.character(player)
  ))

mp_reg_data <- 
  mp_data %>%
  # Filter data for all situations (5v5, player-down, player-up)
  filter(situation == 'all', szn != '12-13') %>%
  # Remove redundant cols
  select(-player_id, -team, -season, -position, -games_played, 
         -penality_minutes, -penalties, -situation, -goals) %>%
  # Transform all names to lowercase
  mutate(name = str_to_lower(name)) %>%
  # Fix nicknames
  mutate(name = case_when(
    str_detect(name, 'tom mccollum') ~ str_replace(name, 'tom mccollum', 'thomas mccollum'),
    str_detect(name, 'j.f. berube') ~ str_replace(name, 'j.f. berube', 'jean-francois berube'),
    str_detect(name, 'j-f berube') ~ str_replace(name, 'j-f berube', 'jean-francois berube'),
    str_detect(name, 'cal petersen') ~ str_replace(name, 'cal petersen', 'calvin petersen'),
    TRUE ~ as.character(name)
  ))

# Join all data together
joined_data <-
  hr_reg_data %>%
  full_join(mp_reg_data, by = c('player' = 'name', 'szn' = 'szn')) %>%
  inner_join(cf_reg_data, by = c('player', 'szn')) %>%
  # Remove extra contract info (since we don't "know" this information yet)
  select(-rk, -cap_hit_percent, -salary, -length, -cap_hit) %>%
  # Same variable as "min" (minutes played)
  select(-icetime)

# Final prep of data
reg_data <- 
  joined_data %>%
  # Code binary variables
  mutate(handed = ifelse(handed == 'Left', 0, 1),
         expiry = ifelse(expiry == 'UFA', 0, 1)) %>%
  # Filter out players whos contracts we want to predict
  filter(!player %in% c('sergei bobrovsky', 'mikko koskinen')) %>%
  # Select numeric predictors
  select_if(is.numeric)

reg_data %>%
  kable() %>%
  kable_styling(bootstrap_options = c('condensed', 'hover')) %>%
  row_spec(0, background = '#e6f0ff') %>%
  scroll_box(width = '100%', height = '250px')
```
  
  
We can see that there are *way* too many predictors (`r reg_data %>% ncol()` columns). So before we build our model, we want to conduct feature selection to reduce the number of predictors in our model. 

#### Feature selection

To do the feature selection, we calculated the information gain for each predictor with response variable AAV, then subsetted for the 10 predictors with the highest amount of information gain. 

```{r, echo = FALSE}
## Feature selection
info_gain <- information.gain(formula = aav ~ ., data = reg_data)
subset_info_gain <- cutoff.k(info_gain, k = 10)

feat_selected <- 
  reg_data %>% 
  select(subset_info_gain, aav)

info_gain %>% 
  add_column(attribute = rownames(info_gain)) %>%
  select(attribute, attr_importance) %>%
  top_n(10, attr_importance) %>%
  arrange(desc(attr_importance)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c('condensed', 'hover'), full_width = F) %>%
  row_spec(0, background = '#e6f0ff')
```

Now that we have the 10 variables with the most information gain, we can use these variables to build our model and then evaluate.  

```{r linear model}
## Model
mod <- lm(aav~., feat_selected)

# Coefficient summary
mod %>% 
  tidy() %>%
  kable() %>%
  kable_styling(bootstrap_options = c('condensed', 'hover')) %>%
  row_spec(0, background = '#e6f0ff')

# Full model summary
mod %>%
  glance() %>%
  select(1:6) %>%
  kable() %>%
  kable_styling(bootstrap_options = c('condensed', 'hover')) %>%
  row_spec(0, background = '#e6f0ff')
```

Looking at our model summary, we see that the only variables that are significant at a level of significance of $\alpha$ = 0.05 other than the model intercept are ```x_play_continued_in_zone``` (expected play continued in the goalies' zone after a shot is taken) and ```low_danger_xgoals``` (expected goals in low danger situations), and ```x_on_goal``` (expected on goal shots in all situations). We are interested in running the model with just these predictors to see if we can get a better fit (by looking at the ```adj.r.squared``` value).

```{r single predictor}
## Model
mod2 <- lm(aav ~ x_play_continued_in_zone + low_dangerx_goals + x_on_goal, data = feat_selected)

# Coefficient summary
mod2 %>% 
  tidy() %>%
  kable() %>%
  kable_styling(bootstrap_options = c('condensed', 'hover')) %>%
  row_spec(0, background = '#e6f0ff')

# Full model summary
mod2 %>%
  glance() %>%
  select(1:6) %>%
  kable() %>%
  kable_styling(bootstrap_options = c('condensed', 'hover')) %>%
  row_spec(0, background = '#e6f0ff')
```

The adjusted $R^2$ value for this single predictor model is *slightly* smaller than for the full 10 predictor model, so we say that the other variables add to the fit of the model, and will use the full model to make our predictions.  

### Predictions

Our model assesses AAV on an individual season basis, so when doing predictions we have to select which season we want to base our predictions on. The goalies we are interested in investigating are two high-profile goalie contracts that were signed last year: Sergei Bobrovsky and Mikko Koskinen.  

#### Sergei Bobrovsky

Bobrovsky got paid $10M AAV when he signed his new contract for this season (until 2026). Assuming we were evaluting Bobrosky in a vaccumm of only his most recent season statistics (2018-19), what does our model say his AAV should have been?

```{r}
bob_data <- 
  joined_data %>%
  filter(player == 'sergei bobrovsky', szn == '18-19') %>%
  select(subset_info_gain)

tibble(predicted = predict(mod, bob_data)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c('condensed', 'hover'), full_width = F) %>%
  row_spec(0, background = '#e6f0ff')
```

This is *significantly* less than what Bobrovsky was actually making in the 2018-19 season.  

#### Mikko Koskinen

Mikko Koskinen was a hot-topic goalie last year because he played less than 20 NHL games in total before the general manager of the Edmonton Oilers (Koskiken's team) signed him to an extension of \$4.5M AAV (effective 2019-20 season) over 3 years. As we've seen in this report, \$4.5M was significantly over the average salary for a goalie in the NHL last season. This contract was heavily critisized because there wasn't an adequate sample size to gauge Koskinen on and there was no rush to sign him. Peter Chiarelli, the GM of the Edmonton Oilers at the time, was already under fire for his many other poor signings and was fired less than 2 weeks after this signing in particular.  

We're curious to see if our model agrees with how much Chiarelli valued Koskinen at, based on Koskinen's full season performance after he signed the contract.    

```{r}
kosk_data <-
  joined_data %>%
  filter(player == 'mikko koskinen') %>%
  select(subset_info_gain)

tibble(predicted = predict(mod, kosk_data)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c('condensed', 'hover'), full_width = F) %>%
  row_spec(0, background = '#e6f0ff')
```

Our model actually values Koskinen at \$3.5M, \$1M less than what Chiarelli signed him for.  


## Conclusions
***

Goalie contracts are some of the hardest to evaluate in hockey. Some goalies get paid a lot for being the most important player on their team, however some get paid way too much just to underperform. Some goalies are playing really well and aren't making what they're worth, but with the way we can see contracts inflating, especially for elite goaltending, they are likely to get big pay days in the coming seasons, as well as drive up the market-value for goalies.  

It is possible that goalies may become the highest paid players in the NHL in the coming years since teams seem to be more willing to shell-out for decent goaltending, as it becomes clearer that goalies contribute in a large way to their teams success or failure.  


## Data Sources
***
Find all data and code hosted in this project's public [Github repository](https://github.com/MaiaPelletier/nhlgoalieevaluation/).   

- HockeyReference (https://www.hockey-reference.com/leagues/NHL_2019_goalies.html#stats::none)
- CapFriendly (https://www.capfriendly.com/browse/active/2019/caphit/all/goalies)
- MoneyPuck (http://moneypuck.com/data.htm)
- MetaHockey (http://metahockey.com/resources/data/sources/compiled/)

## References
***

### Report formatting  
- Side navigation menu (https://bookdown.org/yihui/rmarkdown/html-document.html)
- Theme (https://www.datadreaming.org/post/r-markdown-theme-gallery/)
- Horizontal lines (https://holtzy.github.io/Pimp-my-rmd/#horizontal_lines) 
- Custom CSS (https://bookdown.org/yihui/rmarkdown/custom-css-1.html)
- Tables (project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#table_styles)
- Links (https://rmarkdown.rstudio.com/authoring_basics.html)

### Graphics
- https://github.com/ropensci/rnaturalearth
- https://github.com/nschiett/fishualize
- https://www.r-graph-gallery.com/bubble_chart_interactive_ggplotly.html
- https://stackoverflow.com/questions/36325154/how-to-choose-variable-to-display-in-tooltip-when-using-ggplotly
- https://ggplot2.tidyverse.org/reference/geom_rug.html

### Goaltenders and Goaltending 
- https://www.omha.net/news_article/show/631894-the-evolution-of-different-goaltending-styles
- https://www.sportsnet.ca/hockey/nhl/oilers-koskinen-contract-extension-questions/
- https://www.sportsnet.ca/hockey/nhl/panthers-sign-sergei-bobrovsky-seven-year-70m-deal/

## Appendix
***
```{r}
sessionInfo()
```