---
title: "NHL Goalie Contract Evaluation"
output: ioslides_presentation
author: Ange Michaella Niyonkuru, Erhardt Jansen van Rensburg, and Maia Pelletier
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# slides customization
# https://bookdown.org/yihui/rmarkdown/slide-backgrounds.html

## Load libraries
library(tidyverse)
library(teamcolors)
library(ggsci)
library(rnaturalearth)
library(sf)
library(sp)
library(fishualize)
library(plotly)
library(ggrepel)
library(patchwork)
library(janitor)
library(readxl)
library(lubridate)
library(ggthemes)
library(stringi)
library(FSelector)
library(kableExtra)

theme_set(theme_minimal())

path <- 'data/'

# Combine MetaHockey ------------------------------------------------------

mh_files <- paste0(path, list.files(path, pattern = 'MH_'))

## General cleaning function
clean_mh_data <- function(mh_data) {
  
  tidy_mh_data <- 
    mh_data %>%
    clean_names() %>%
    mutate(cap_hit = as.numeric(cap_hit)) %>%
    mutate(dob = date(parse_date_time(dob, orders = c('mdy', 'ymd'))))
  
  if (any(str_detect(names(tidy_mh_data), 'chip'))) {
    tidy_mh_data <- 
      tidy_mh_data %>%
      mutate(chip = as.numeric(chip))
  }
  
  if(any(str_detect(names(tidy_mh_data), 'dist'))) {
    tidy_mh_data <- 
      tidy_mh_data %>%
      mutate(dist = as.character(dist))
  }
  
  return(tidy_mh_data)
}


mh_szns <- paste0(12:17, '-', 13:18)

## Read/clean data and append to list
mh_data_list <- list()
for (i in 1:length(mh_files)) {
  mh_data_list[[i]] <- 
    read_excel(mh_files[i]) %>%
    clean_mh_data() %>%
    mutate(szn = mh_szns[i])
}

## Dataset specific cleaning
mh_data_list[[2]] <- 
  mh_data_list[[2]] %>% 
  rename(`gp` = gp_16, gs = gs_17, sa = sa_21, sv_percent = sv_percent_24)

mh_data_list[[3]] <- 
  mh_data_list[[3]] %>%
  rename(gp = gp_17)

## Combine data into one dataset
new_mh_data <- tibble()
for (i in 1:length(mh_data_list)) {
  new_mh_data <- bind_rows(new_mh_data, mh_data_list[[i]])
}

## Clean dataset
getblanks <- function(blanks) {
  
  blank <- sum(is.na(new_mh_data[blanks]))
  column_name <- colnames(new_mh_data[blanks])
  ans <- tibble(column_id = blanks, column_name = column_name,  NA_amount = blank)
  
  return(ans)
}

## Find number of blanks
response <- tibble()
for (i in 1:ncol(new_mh_data)) {
  response  <- bind_rows(response, getblanks(i))
}

## Remove column if filled with blanks (no match in other datasets)
for (y in 1:nrow(response)) {
  
  if (response[y, "NA_amount"] > 81) {
    
    name_column <- paste((response[y,"column_name"]))
    
    new_mh_data <- 
      new_mh_data %>% 
      select(-name_column)
  }
}

mh_data <- 
  new_mh_data %>%
    unite('player', c('first_name', 'last_name'), sep = ' ')

# Combine HockeyReference -------------------------------------------------

hr_files <- paste0(path, list.files(path, pattern = 'HR_'))

num_of_sheets <- length(excel_sheets(hr_files))

hr_szns <- paste0(12:18, '-', 13:19)

# General cleaning function
clean_hr_data <- function(hr_data) {
  
  hr_data <- 
    hr_data %>%
    clean_names() %>%
    mutate(szn = hr_szns[i]) %>%
    mutate(player = str_remove(player, '\\\\[\\w]+')) %>%
    mutate(player = str_remove(player, '\\*')) %>%
    add_count(player, name = 'num_of_teams') %>%
    mutate(
      num_of_teams = ifelse(num_of_teams > 1, num_of_teams - 1, num_of_teams)
    )
  
  clean_data <-
    bind_rows(
      ## Traded players
      hr_data %>%
        filter(num_of_teams > 1) %>%
        filter(tm == 'TOT'),
      ## Players that played on the same team all szn
      hr_data %>%
        filter(num_of_teams == 1)
    ) %>%
    rename(team = tm) %>%
    arrange(rk)
  
  
  return(clean_data)
}

## Append data from each sheet in excel file to empty list and clean
hr_data_list <- list()
for (i in 1:num_of_sheets) {
  hr_data_list[[i]] <- 
    read_excel(hr_files, sheet = i, skip = 1) %>% 
    clean_hr_data()
}

## Combine data into on dataset
hr_data <- tibble()
for (i in 1:length(hr_data_list)) {
  hr_data <- 
    bind_rows(hr_data, hr_data_list[[i]])
}


# Combine MoneyPuck -------------------------------------------------------

mp_files <- paste0(path, list.files(path, pattern = 'MP_'))

mp_data_list <- list()
for(i in 1:length(mp_files)) {
  mp_data_list[[i]] <- 
    read_csv(mp_files[i]) %>%
      clean_names() %>%
      mutate(szn = hr_szns[i])
}

## Note: MoneyPuck data was perfectly clean (that we can tell) so cleaning was
## not needed before creating dataset

mp_data <- tibble()
for (i in 1:length(mp_data_list)) {
  mp_data <- bind_rows(mp_data, mp_data_list[[i]])
}

# Combine CapFriendly -------------------------------------------------

cf_files <- paste0(path, list.files(path, pattern = 'CF-'))

num_of_sheets <- length(excel_sheets(cf_files))

cf_szns <- excel_sheets(cf_files)

## Append data from each sheet in excel file to empty list
cf_data_list <- list()
for (i in 1:num_of_sheets) {
  cf_data_list[[i]] <- 
    read_excel(cf_files, sheet = i) %>% 
    clean_names() %>%
    mutate(szn = cf_szns[i])
}

## Small tidying
for (i in 1:length(cf_data_list)) {
  cf_data_list[[i]] <-
    cf_data_list[[i]] %>%
    # When a player dies(during the szn?), it denotes "Died" in Age column
    mutate(age = str_extract(age, '\\d+'),
           age = as.numeric(age)) %>%
    # The players column contains a numbered index (i.e. 1. Carey Price)
    mutate(player = str_remove(player, '[0-9]*\\.\\s*'))
}
  
## Combine data into on dataset
cf_data <- tibble()
for (i in 1:length(cf_data_list)) {
  cf_data <- 
    bind_rows(cf_data, cf_data_list[[i]]) 
}

cf_data <-
  cf_data %>%
    separate(drafted, into = c('overall', 'draft_round', 'draft_year'), sep = ' - ') %>%
    separate(draft_year, into = c('draft_year', 'draft_team'), sep = ' ') %>%
    mutate(overall = as.numeric(overall),
           draft_year = as.numeric(draft_year),
           draft_team = str_remove_all(draft_team, '[\\( \\)]'))

# Combine NHL Standings ---------------------------------------------------

standings_files <- paste0(path, list.files(path, pattern = 'nhl_standings'))

num_of_sheets <- length(excel_sheets(standings_files))

standings_szns <- paste0(12:18, '-', 13:19)

## Combine data into one dataset
standings_data <- tibble()
for (i in 1:num_of_sheets) {
  add <- 
    read_excel(standings_files, sheet = i, skip = 1) %>% 
    clean_names() %>%
    mutate(szn = standings_szns[i]) %>%
    select(rk, team = x2, w, l, ol, pts, szn) %>%
    mutate(team = str_remove(team, '\\*'))
  
  standings_data <- bind_rows(standings_data, add)
}


```

## Objectives

- How good are goalies?
- Do they work hard for the money?
- How much to pay goalies?

## Best goalies (& worst goalies) {data-background=#ebf5ff}

- How many goalies are we looking at?
- Where are the goalies from and what do they look like? 
- Who are the best ones?
- Top 10 goalies over the past 6 seasons
- Worst 10 goalies over the last 6 seasons
- Top goalies by season

## Goalie overview

- How many goalies are we looking at total in the last 7 seasons?
 
```{r Total number of goalies}
hr_data %>% 
  distinct(player) %>% 
  nrow()
```

- How many games do goalies play in general?
```{r Avg num of GP, echo = FALSE}
hr_data %>%
  summarise(avg_no_games = mean(gp)) %>%
  pull(avg_no_games)
```

## How many games do goalies play in general?
```{r GP Histogram, fig.height=3}
ggplot(hr_data, aes(gp)) +
  geom_histogram(binwidth = 5, fill = '#a7d2cb', alpha = 0.8) +
  labs(x = 'Games played in the NHL by goalies', y = 'Frequency',
       caption = '1 Game played = 1 or more minute played on ice') +
  theme(plot.caption = element_text(color = 'grey50'))
```

- How many of those goalies *actually* played in the the NHL? (>40 games played)

```{r GP over 40}
hr_data %>%
  # 40 games ~ just under a full szn for a goalie
  filter(gp > 40) %>%
  distinct(player) %>%
  nrow()
```


## Goalie demographics

- Where are NHL goalies from?

```{r Map, warning = FALSE, message = FALSE}

## Plot map of where goalies are from

goalie_countries <- 
  cf_data %>%
    count(country, name = "num_of_goalies") %>%
    na.omit()

world <- 
  ne_countries(
    scale = "medium", 
    returnclass = "sf", 
    continent = c('North America', 'Europe')
  )

world %>%
  left_join(goalie_countries, by = c('name' = 'country')) %>%
  ggplot() +
  geom_sf(aes(fill = num_of_goalies)) +
  labs(fill = "Number of goalies") + 
  scale_fill_fish(option = "Prionace_glauca", direction = -1) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.key.height = unit(2, 'mm'),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8))

```

## Goalie demographics

- What do they look like (height & weight)?

```{r Height and weight, warning = FALSE}

cf_data_player_height_weight <- cf_data %>% 
  select(player, weight, height, age, country) %>%
  clean_names() %>%
  unique() %>% 
  mutate(weight_kg = sub(".*-","", weight) %>% parse_number(), 
         height_cm = sub(".*-","",height) %>% parse_number(), 
         text = paste("Player: ", player, "\nAge: " , age, "\nCountry: ", country)) 

p <- cf_data_player_height_weight %>%
  ggplot(aes(x = weight_kg, y = height_cm, colour = age)) + 
  geom_point(aes(label = player, label2 = country), alpha = 0.7) + 
  labs(
    x = "Player weight in kg",
    y= "Player height in cm",
    title = "Goalie height and weight chart",
    colour = "Player age"
  ) +
  scale_color_fish(option = "Prionace_glauca", direction = -1)

ggplotly(p)

```


## Best goalies 

```{r, warning = FALSE}

hr_data <- hr_data %>% mutate(w_percent = w/gp)

Salaries <- cf_data %>% 
  group_by(player) %>% 
  summarise(mean_aav = mean(aav)) %>% 
  mutate(player = replace(player,player == "Marc-André Fleury","Marc-Andre Fleury")) %>%
  mutate(player = replace(player,player == "Jaroslav Halák","Jaroslav Halak")) %>% 
  mutate(player = replace(player, player == "Eddie Läck","Eddie Lack")) %>% 
  mutate(player = replace(player, player == "Jacob Markström","Jacob Markstrom")) %>%
  mutate(player = replace(player, player == "Petr Mrázek","Petr Mrazek"))

Best_Goalies <- 
  hr_data %>% 
  filter(!(player == "Martin Jones" & szn == "14-15")) %>%
  group_by(player) %>% 
  summarise(gp = sum(gp),
            w = sum(w),
            ga = sum(ga),
            sa = sum(sa),
            sv = sum(sv),
            ) %>%
  filter(gp > 120) %>% 
  mutate(mean_w_percent = w/gp, 
         mean_sv_percent = sv/sa,
         mean_gaa = ga/gp,
         avg_sv = mean(mean_sv_percent),
         mean_gsaa = (sa * (1-avg_sv))- ga,
         )

Best_Goalies <- Best_Goalies %>% 
  left_join(Salaries, by = c("player" = "player"))

Mean_sv_plot <- 
  Best_Goalies %>% 
  top_n(10, mean_sv_percent) %>%
  ggplot(aes(x = mean_sv_percent, y = mean_aav)) + 
  geom_point(size = 2, color = '#a7d2cb') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Save Percentage", y = "Average Annual Salary")

Mean_gaa_plot <- 
  Best_Goalies %>% 
  top_n(-10, mean_gaa) %>%
  ggplot(aes(x = mean_gaa, y = mean_aav)) + 
  geom_point(size = 2, color = '#f2d388') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Against Average", y = "Average Annual Salary")

mean_w_percent_plot <- 
  Best_Goalies %>% 
  top_n(10, mean_w_percent) %>% 
  ggplot(aes(x = mean_w_percent, y = mean_aav)) + 
  geom_point(size = 2, color = '#c98474') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Percentage of Games Won", y = "Average Annual Salary")


mean_gsaa_plot <- 
  Best_Goalies %>% 
  top_n(10, mean_gsaa) %>% 
  ggplot(aes(x = mean_gsaa,y = mean_aav))+ 
  geom_point(size = 2, color = '#874c62') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Saved Above Average", y = "Average Annual Salary")


best_average_plots <- (Mean_sv_plot | Mean_gaa_plot) / (mean_w_percent_plot | mean_gsaa_plot)
 
best_average_plots + 
  plot_annotation(
    title = "Top 10 Averages By Goalies Over The Last Six Seasons"
    )
```



## Best goalies 

- Top goalies by season

```{r, warning = FALSE}

# Finding the best goalies by Season
# creates tables for each stat we are looking at

top_goalie_sv <- function(season){
  top_goalie_stats <- hr_data %>% filter(szn == season, gp > 40) 
  top_goalie_sv_percent <- top_goalie_stats %>% top_n(1,sv_percent) %>%
    select(player, sv_percent, szn)
  return(top_goalie_sv_percent)
}

top_goalie_gaa <- function(season){
  top_goalie_stats <- hr_data %>% filter(szn == season, gp > 40)
  goalie_gaa <- top_goalie_stats %>% top_n(-1,gaa) %>% 
    select(player, gaa, szn)
  return(goalie_gaa)
}


top_goalie_w_percent <- function(season){
  top_goalie_stats <- hr_data %>% filter(szn == season, gp > 40)
  goalie_w_percent <- top_goalie_stats %>% top_n(1,w_percent) %>% 
    select(player, w_percent, szn)
  return(goalie_w_percent)
}


top_goalie_gsaa <- function(season){
  top_goalie_stats <- hr_data %>% filter(szn == season, gp > 40)
  goalie_gsaa <- top_goalie_stats %>% top_n(1,gsaa) %>% 
    select(player, gsaa, szn)
  return(goalie_gsaa)
}

#Loop to fill in the tables

best_sv_year <- tibble()
best_gaa_year <- tibble()
best_win_percent_year <- tibble()
best_gsaa_year <- tibble()
for (i in 1:7){
  best_sv_year <- bind_rows(best_sv_year,top_goalie_sv(hr_szns[i]))
  best_gaa_year <- bind_rows(best_gaa_year,top_goalie_gaa(hr_szns[i]))
  best_win_percent_year <- bind_rows(best_win_percent_year,top_goalie_w_percent(hr_szns[i]))
  best_gsaa_year <- bind_rows(best_gsaa_year,top_goalie_gsaa(hr_szns[i]))
}



#Plots to see best goalies by season 


sv_plot <- 
  best_sv_year %>% 
  ggplot(aes(x = szn, y = sv_percent)) + 
  geom_point(size = 2, color = '#a7d2cb') +
  geom_text_repel(aes(label = player), size = 3) +
  coord_flip() + 
  labs(x = "Season", y = "Save Percentage")

gaa_plot <- 
  best_gaa_year %>% 
  ggplot(aes(x =szn, y = gaa)) + 
  geom_point(size = 2, color = '#f2d388') +
  geom_text_repel(aes(label = player), size = 3) +
  coord_flip() + 
  labs(x = "Season", y = "Goals Against Average")

w_percent_plot <- 
  best_win_percent_year %>% 
  ggplot(aes(x = szn, y = w_percent)) + 
  geom_point(size = 2, color = '#c98474') + 
  geom_text_repel(aes(label = player), size = 3) +
  coord_flip() + 
  labs(x = "Season", y = "Win Percentage")

gsaa_plot <- 
  best_gsaa_year %>% 
  ggplot(aes(x = szn, y = gsaa)) + 
  geom_point(size = 2, color = '#874c62') + 
  geom_text_repel(aes(label = player), size = 3) +
  coord_flip() + 
  labs(x = "Season", y = "Goals Saved Above Average")

year_plots <- (sv_plot | gaa_plot) / (w_percent_plot | gsaa_plot)

year_plots + plot_annotation(
  title = "Top Goalies by Season"
)

```


## Worst Goalies

```{r, warning = FALSE, message = FALSE}

#Worst Goalies by averages overall

worst_mean_sv_plot <- Best_Goalies %>% 
  top_n(-10,mean_sv_percent) %>%
  ggplot(aes(x = mean_sv_percent, y = mean_aav)) + 
  geom_point(size = 2, color = '#a7d2cb') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Save Percentage", y = "Average Annual Salary")

worst_mean_gaa_plot <- Best_Goalies %>% 
  top_n(10,mean_gaa) %>%
  ggplot(aes(x = mean_gaa, y = mean_aav)) + 
  geom_point(size = 2, color = '#f2d388') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Against Average", y = "Average Annual Salary")

worst_mean_w_percent_plot <- Best_Goalies %>% 
  top_n(-10, mean_w_percent) %>% 
  ggplot(aes(x = mean_w_percent, y = mean_aav)) + 
  geom_point(size = 2, color = '#c98474') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Percentage of Games Won", y = "Average Annual Salary")


worst_mean_gsaa_plot <- Best_Goalies %>% 
  top_n(-10, mean_gsaa) %>% 
  ggplot(aes(x = mean_gsaa,y = mean_aav))+ 
  geom_point(size = 2, color = '#874c62') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Saved Above Average", y = "Average Annual Salary")


Worst_average_plots <- (worst_mean_sv_plot | worst_mean_gaa_plot) / (worst_mean_w_percent_plot | worst_mean_gsaa_plot)

Worst_average_plots + plot_annotation(
  title = "Worst 10 Averages By Goalies Over The Last Six Seasons"
)

```


## Worst goalies over last 6 seasons

```{r, warning = FALSE, message = FALSE}

# Some really bad goalies by year 

missing_salary <- 
  read_xls("data/MH_nhl_goalies_2017-2018.xls") %>% 
  clean_names() %>% 
  unite('player', c('first_name', 'last_name'), sep = ' ') %>% 
  filter(player == "Scott Darling") %>% 
  select(player, salary)

worst_goalies <- hr_data %>%
  filter(gp > 40 & sv_percent < 0.9)

worst_salaries <- cf_data %>% select(player, aav, szn)

worst_goalies <- worst_goalies %>% 
  left_join(worst_salaries,by = c("player" = "player", "szn" = "szn")) %>%
  left_join(missing_salary, by = c("player" = "player")) %>% 
  mutate(aav = replace_na(aav,0)) %>% 
  mutate(salary = replace_na(salary,0)) %>% 
  mutate(aav = aav+ salary) %>% 
  unite('player_szn', c('player','szn'), sep = ' | Season:', remove = FALSE) %>%
  select(-salary) %>%
  mutate(w_percent = w/gp)

wg_sv_percent <- worst_goalies %>% 
  ggplot(aes(x = sv_percent, y = aav)) + 
  geom_point(size = 2, color = '#a7d2cb') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Save Percentage", y = "Annual Average Salary")

wg_gsaa <- worst_goalies %>% ggplot(aes(x = gsaa, y = aav)) + 
  geom_point(size = 2, color = '#f2d388') +
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Saved Above Average", y = "Annual Average Salary")

wg_gaa <- worst_goalies %>% ggplot(aes(x = gaa, y = aav)) + 
  geom_point(size = 2, color = '#c98474') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Goals Against Average", y = "Annual Average Salary")

wg_w_percent <- worst_goalies %>% 
  ggplot(aes(x = w_percent, y = aav)) + 
  geom_point(size = 2, color = '#874c62') + 
  geom_text_repel(aes(label = player), size = 3) +
  labs(x = "Win Percentage", y = "Annual Average Salary")

Worst_goalies_plot <- (wg_sv_percent | wg_gaa) / (wg_w_percent | wg_gsaa)

Worst_goalies_plot + plot_annotation(
  title = "Worst perfoming goalies and how much they get paid"
)

```


## Goalies contracts {data-background=#ebf5ff}

- How much goalies get paid in general and how has that changed over time? 
- Does this depend on how good their team is?

## Goalies contracts

- How much do goalies get paid and how has that changed over time? 

```{r, warning = FALSE}

cf_data %>%
  filter(gp > 40) %>%
  mutate(szn = paste0(20, szn)) %>%
  ggplot(aes(szn, aav, fill = szn, color = szn)) +
  geom_violin(alpha = 0.8, show.legend = F, trim = T, adjust = 0.75, width = 1) +
  geom_boxplot(width = 0.1, color = 'grey80', fill = 'grey50', alpha = 0.1) +
  labs(x = NULL, y = 'Average annual value ($)') +
  #ggtitle('Distribution of goalie average annual value') +
  ggsci::scale_color_futurama() +
  ggsci::scale_fill_futurama() +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text.x = element_text(size = 12, color = 'grey25'),
        axis.text.y = element_text(size = 7, color = 'grey25'),
        axis.title.y = element_text(size = 9, color = 'grey25', face = 'italic'),
        plot.title = element_text(hjust = 0.5, size = 16, color = 'grey15', face= 'italic'))
```

## Contract inflation

- How are goalie contracts inflating?

```{r Contract inflation, warning=FALSE, message=FALSE}
cf_data %>%
  group_by(exp_year) %>%
  summarise(med_aav = median(aav)) %>%
  ggplot(aes(exp_year, med_aav)) +
  geom_step(color = 'turquoise3', size = 1) +
  labs(x = 'Contract expiry year', y = 'Median average annual value of contract') +
  #scale_fill_fish(option = "Prionace_glauca", direction = -1) +
  theme_classic() +
  theme(axis.line = element_blank())
```
## Goalies contracts
 
- Who are the Top 10 paid players in the last 7 years?

```{r, warning = FALSE}

## Get top 10 paid players in league over last 7 years
top_10_paid <- 
  cf_data %>%
    group_by(player) %>%
    top_n(1, aav) %>%
    ungroup() %>%
    distinct(player, aav, team) %>%
    top_n(10, aav)

## Plot top 10 paid players
top_10_paid %>%
  mutate(point_lab = paste0('$', as.integer(aav))) %>%
  mutate(team = substr(team, start = str_length(team) - 2, stop = str_length(team))) %>%
  mutate(player = paste0(player, '\n(', team, ')     ')) %>%
  left_join(teamcolors, by = c('team' = 'name')) %>%
  ggplot() +
  geom_segment(aes(x = reorder(player, aav), xend = player, y = 0, yend = aav), color = 'grey50') +
  geom_point(aes(x = player, y = aav, color = team), size = 5, show.legend = F) +
  labs(x = NULL, y = 'AAV($)') +
  #ggtitle('Top 10 paid players in the league (as of 2018-19)') +
  coord_flip() +
  scale_color_futurama() +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text.y = element_text(size = 10, color = 'grey25'),
        plot.title = element_text(hjust = 0.5, size = 16, color = 'grey25', face= 'italic'))

```

## Goalies contracts
 
- Do this depend on how good their team is?

```{r}
league_avgs <-
  hr_data %>%
  filter(gp > 20) %>%
  group_by(szn) %>%
  mutate(w_percent = w/gp) %>%
  summarise(
    avg_sv_percent = mean(sv_percent),
    avg_gaa = mean(gaa),
    avg_w_percent = mean(w_percent),
    avg_gsaa = mean(gsaa)
  )

starting_goalies_best <- 
  standings_data %>%
    filter(rk == 1) %>%
    select(-w, -l, -ol, -pts) %>%
    left_join(read_csv('data/NHLTeams.csv')) %>%
    left_join(hr_data, by = c('abbrev' = 'team', 'szn' = 'szn')) %>%
    group_by(team, szn) %>%
    top_n(1, gp) %>%
    mutate(w_percent = w/gp)

starting_goalies_sv_percent <- 
  ggplot() +
    geom_point(data = starting_goalies_best, 
               aes(x = szn, y = sv_percent, color = team), 
               show.legend = F, size = 4) +
    geom_point(data = league_avgs,
               aes(x = szn, y = avg_sv_percent),
               show.legend = F, size = 3, color = 'red3') +
    geom_text_repel(data = starting_goalies_best, 
                    aes(x = szn, y = sv_percent, label = player), 
                    size = 3) +
    labs(y = 'Save percent', x = 'Season') +
    ggtitle('Save percent') +
    scale_color_fish_d(option = "Prionace_glauca") +
    coord_flip() +
    theme_hc()

starting_goalies_gaa <- 
  ggplot() +
    geom_point(data = starting_goalies_best, 
               aes(x = szn, y = gaa, color = team), 
               show.legend = F, size = 4) +
    geom_point(data = league_avgs,
               aes(x = szn, y = avg_gaa),
               show.legend = F, size = 3, color = 'red3') +
    geom_text_repel(data = starting_goalies_best, 
                    aes(x = szn, y = gaa, label = player), 
                    size = 3) +
    labs(y = 'Goals against avg', x = NULL) +
    ggtitle('Goals against average') +
    scale_color_fish_d(option = "Prionace_glauca") +
    coord_flip() +
    theme_hc()

starting_goalies_comb <- starting_goalies_sv_percent / starting_goalies_gaa
```

## Can we predict how much goalies with get paid? {data-background=#ebf5ff}

- Build multivariate regression model to predict contract average annual value

```{r Regression prep}
## Prep and combine data

cf_reg_data <- 
  cf_data %>%
  mutate(
    ## Removes accents off names
    player = stri_trans_general(player, 'latin-ascii'),
    ## Transform all names to lowercase
    player = str_to_lower(player)) %>%
  ## Remove redundant cols
  select(-weight, -height, -pos, -team, -age, -gp, -w, -l, -so, -gaa, -sv_percent)

hr_reg_data <-
  hr_data %>%
  ## Transform all names to lowercase
  mutate(player = str_to_lower(player)) %>%
  # Filter for post lockout szns
  filter(szn != '12-13') %>%
  # Fix nicknames
  mutate(player = case_when(
    str_detect(player, 'cal heeter') ~ str_replace(player, 'cal heeter', 'calvin heeter'),
    str_detect(player, 'matt o\'connor') ~ str_replace(player, 'matt o\'connor', 'matthew o\'connor'),
    str_detect(player, 'eddie pasquale') ~ str_replace(player, 'eddie pasquale', 'edward pasquale'),
    TRUE ~ as.character(player)
  ))

mp_reg_data <- 
  mp_data %>%
  # Filter data for all situations (5v5, player-down, player-up)
  filter(situation == 'all', szn != '12-13') %>%
  # Remove redundant cols
  select(-player_id, -team, -season, -position, -games_played, 
         -penality_minutes, -penalties, -situation, -goals) %>%
  # Transform all names to lowercase
  mutate(name = str_to_lower(name)) %>%
  # Fix nicknames
  mutate(name = case_when(
    str_detect(name, 'tom mccollum') ~ str_replace(name, 'tom mccollum', 'thomas mccollum'),
    str_detect(name, 'j.f. berube') ~ str_replace(name, 'j.f. berube', 'jean-francois berube'),
    str_detect(name, 'j-f berube') ~ str_replace(name, 'j-f berube', 'jean-francois berube'),
    str_detect(name, 'cal petersen') ~ str_replace(name, 'cal petersen', 'calvin petersen'),
    TRUE ~ as.character(name)
  ))

# Join all data together
joined_data <-
  hr_reg_data %>%
  full_join(mp_reg_data, by = c('player' = 'name', 'szn' = 'szn')) %>%
  inner_join(cf_reg_data, by = c('player', 'szn')) %>%
  # Remove extra contract info (since we don't "know" this information yet)
  select(-rk, -cap_hit_percent, -salary, -length, -cap_hit) 

reg_data <- 
  joined_data %>%
  # Code binary variables
  mutate(handed = ifelse(handed == 'Left', 0, 1),
         expiry = ifelse(expiry == 'UFA', 0, 1)) %>%
  # Select numeric predictors
  select_if(is.numeric)
```

## Feature selection
- *Way* too many predictors
```{r}
reg_data %>% ncol()
```

- Calculate **information gain** for predictors and grab 10 highest predictors in information gain
```{r, echo = FALSE}
## Feature selection
info_gain <- information.gain(formula = aav ~ ., data = reg_data)
subset_info_gain <- cutoff.k(info_gain, k = 10)

feat_selected <- 
  reg_data %>% 
  select(subset_info_gain, aav)

feat_selected %>% head() %>% kable() %>% kable_styling()
```

## Build model
- Build regression model with selected features

```{r linear model}
## Model
mod <- lm(aav~., feat_selected)
summary(mod)
```

## Predictions

- Sergei Bobrovsky

```{r}
new_data1 <- 
  joined_data %>%
  filter(player == 'sergei bobrovsky', szn == '18-19') %>%
  select(subset_info_gain, aav)

predict(mod, new_data1)
```

- Mikko Koskinen 

```{r}
new_data2 <-
  joined_data %>%
  filter(player == 'mikko koskinen') %>%
  select(subset_info_gain, aav)

predict(mod, new_data2)
```


## Conclusions 

## Thank you
- Questions?


